---
title: "nba6340-hw2"
author: "Justin Ngai"
date: "2023-02-15"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Setup

Install and load libraries
```{r warning=FALSE, message=FALSE, echo=FALSE}
library(tidyverse)
library(readxl)
library(skimr)
library(ggplot2)
library(dplyr)
library(fmsb)
```

We first load the excel data and gain some basic insights on the distribution of the data
```{r echo=TRUE, message=FALSE}
malm <- read_excel("MALM.xlsx", sheet="MALM", col_names = TRUE, col_types = NULL)

# Convert "Buyer" column to bool
stopifnot(length(unique(malm$buyer)) == 2) # verify that there are only "1" and "0" values
malm$buyer <- as.logical(malm$buyer)

# Convert "Gender" column to bool
stopifnot(length(unique(malm$buyer)) == 2) # verify that there are only "F" and "M" values
malm <- malm %>% 
  mutate(
    gender = recode(gender, "M"=1, "F"=0) # Male as 1 and Female as 0
  )

head(malm, 10)
dim(malm)

# glimpse(malm)
# summary(malm)

# skim(malm)

```




## Part I: Logistic Regression

**1. Define dependent variable(s) to apply the logistic regression.**

We set `buyer` as the dependent variable as we want to predict whether or not the consumers will ultimately decide to purchase the product. Also since we are working with logistic regression which only supports binary dependent variables, the `buyer` variable is the best choice we have.

**2. Define independent variable(s). **

```{r}
# Specify your own feature variables here
outcome_vars = c("buyer") 

feature_vars = c(
  "gender", # prob not very helpful
  "state", # prob not very helpful
  "zip", # prob not very helpful
  "zip3", # prob not very helpful
  "first",
  "last",
  "meal$",
  "nonmeal$",
  "total$", # just meal + nonmeal
  "purch",  # just the following summed
  "dairy_free",
  "poultry",
  "pork",
  "beef",
  "gluten_free",
  "vegetarian",
  "seafood"
)

xfeature_vars = c(
  "gender",
  "last",
  "total$", # meal$ + nonmeal$
  "purch"   # purch is the sum of all the categories
)
```

```{r}
logreg = glm(buyer ~ 
              gender + 
              # first +
              last +
              `meal$` +
              # `nonmeal$` +
              # `total$` +    # This is just meal$ + nonmeal$
              # purch +       # This is just the sum of the following categories
              # dairy_free +
              # poultry +
              # pork +
              # beef +
              gluten_free +
              vegetarian +
              seafood,
             data=malm, 
             family="binomial")

# the output are the coefficients:
summary(logreg)
NagelkerkeR2(logreg)
```

## Part II: Decile Analysis of Logistic Regression Results

**4. Assigning each customer to a decile based on predicted probability of purchase.**

Add predictions
```{r}
malm$pred <- predict(logreg, malm, type = "response")
```

Rank into deciles
```{r}
malm_d <- malm %>%
  mutate(
    p_rank = ntile(-pred, 10),     # Larger pred value = more likely to buy
  )
```

Plot mean *predicted* response rate per decile
```{r}
malm_d_avg <- malm_d %>% 
  group_by(p_rank) %>% 
  summarise(
    avg_pred = mean(pred),
  )
malm_d_avg

ggplot(malm_d_avg, aes(x=p_rank, y=avg_pred)) + 
  geom_col(fill="#69b3a2", color="#e9ecef", alpha=0.9) + 
  ggtitle('Mean pred value per decile')+
  scale_x_continuous(breaks=c(1,2,3,4,5,6,7,8,9,10))
```

Plot mean real response rate per decile
```{r}
malm_d_avg_real <- malm_d %>% 
  group_by(p_rank) %>% 
  summarise(
    avg_buyer = mean(buyer),
  )
malm_d_avg_real

ggplot(malm_d_avg_real, aes(x=p_rank, y=avg_buyer)) + 
  geom_col(fill="#b3697a", color="#e9ecef", alpha=0.9) + 
  ggtitle('Mean real response rate per decile')+
  scale_x_continuous(breaks=c(1,2,3,4,5,6,7,8,9,10))
```

**5. Generate a report showing the number of customers, the number of buyers of White Truffle Risotto and the response rate to the offer by decile. Summarize and interpret the results.**
```{r}
num_customers = nrow(malm)
num_customers

num_buyers = sum(malm$buyer)
num_buyers

malm_d %>% group_by(p_rank) %>% summarise(
  avg_response_rate = mean(buyer),
  num_customers = n(),
  num_buyers = sum(buyer)
)
```

**6. Generate a report showing the mean values of the independent variables by the probability of purchase decile. **

```{r}
mean_gender <- malm_d %>% group_by(p_rank) %>% summarise(avg_gender=mean(gender))
mean_last <- malm_d %>% group_by(p_rank) %>% summarise(avg_last=mean(last))
mean_meal <- malm_d %>% group_by(p_rank) %>% summarise(avg_meal=mean(`meal$`))
mean_gluten_free <- malm_d %>% group_by(p_rank) %>% summarise(avg_gluten_free=mean(gluten_free))
mean_vegetarian <- malm_d %>% group_by(p_rank) %>% summarise(avg_vegetarian=mean(vegetarian))
mean_seafood <- malm_d %>% group_by(p_rank) %>% summarise(avg_seafood=mean(seafood))

# Gender
ggplot(mean_gender, aes(x=p_rank, y=avg_gender)) + 
  geom_col(fill="#b3697a", color="#e9ecef", alpha=0.9) + 
  ggtitle('Mean of gender by decile')+
  scale_x_continuous(breaks=c(1,2,3,4,5,6,7,8,9,10))

# Last
ggplot(mean_last, aes(x=p_rank, y=avg_last)) + 
  geom_col(fill="#b3697a", color="#e9ecef", alpha=0.9) + 
  ggtitle('Mean of last by decile')+
  scale_x_continuous(breaks=c(1,2,3,4,5,6,7,8,9,10))

# Meal
ggplot(mean_meal, aes(x=p_rank, y=avg_meal)) + 
  geom_col(fill="#b3697a", color="#e9ecef", alpha=0.9) + 
  ggtitle('Mean of meal by decile')+
  scale_x_continuous(breaks=c(1,2,3,4,5,6,7,8,9,10))

# Gluten Free
ggplot(mean_gluten_free, aes(x=p_rank, y=avg_gluten_free)) + 
  geom_col(fill="#b3697a", color="#e9ecef", alpha=0.9) + 
  ggtitle('Mean of gluten free by decile')+
  scale_x_continuous(breaks=c(1,2,3,4,5,6,7,8,9,10))

# Vegetarian
ggplot(mean_vegetarian, aes(x=p_rank, y=avg_vegetarian)) + 
  geom_col(fill="#b3697a", color="#e9ecef", alpha=0.9) + 
  ggtitle('Mean of vegetarian by decile')+
  scale_x_continuous(breaks=c(1,2,3,4,5,6,7,8,9,10))

# Seafood
ggplot(mean_seafood, aes(x=p_rank, y=avg_seafood)) + 
  geom_col(fill="#b3697a", color="#e9ecef", alpha=0.9) + 
  ggtitle('Mean of seafood by decile')+
  scale_x_continuous(breaks=c(1,2,3,4,5,6,7,8,9,10))

```

**7. Use the information from the report above to create a table (and/or chart) showing the lift and cumulative lift for each decile.**

```{r}

#change buyer to 1 true and 0 false
malm_d$buyer <- as.integer(as.logical(malm_d$buyer))

#add variables (lift & gain, response rate) to malm_d_avg_real, 
malm_d_avg_real <- malm_d %>% 
  group_by(p_rank) %>% 
  summarise(
    avg_buyer = mean(buyer),
    customer = length(buyer),
    buyer = sum(buyer),
  )

#make lift & gain table
malm_d_avg_real$cumcustomer <- cumsum(malm_d_avg_real$customer)
malm_d_avg_real$cumbuyer <- cumsum(malm_d_avg_real$buyer)
malm_d_avg_real$resprate <- (malm_d_avg_real$buyer/malm_d_avg_real$customer)
malm_d_avg_real$cumresprate <- (malm_d_avg_real$cumbuyer/malm_d_avg_real$cumcustomer)
malm_d_avg_real$lift <- (malm_d_avg_real$resprate/mean(malm_d_avg_real$resprate)*100)
malm_d_avg_real$cumlift <- (malm_d_avg_real$cumresprate/mean(malm_d_avg_real$resprate)*100)
malm_d_avg_real$gains <- (malm_d_avg_real$buyer/sum(malm_d_avg_real$buyer))
malm_d_avg_real$cumgains <- (cumsum(malm_d_avg_real$gains))

#lift
ggplot(data=malm_d_avg_real, aes(x=p_rank, y=lift, group=1)) +
  geom_line(linetype = "dashed", color = "#69b3a2")+
  ggtitle('Cumulative lift and lift')+
  geom_line(data=malm_d_avg_real,aes(p_rank, y=cumlift), color = "#b3697a", linetype = "twodash")+
  scale_x_continuous(breaks=c(1,2,3,4,5,6,7,8,9,10))

#gains
ggplot(data=malm_d_avg_real, aes(x=p_rank, y=gains, group=1)) +
  geom_line(linetype = "dashed", color = "#69b3a2")+
  ggtitle('Cumulative gains and gains')+
  geom_line(data=malm_d_avg_real,aes(x=p_rank, y=cumgains), color = "#b3697a", linetype = "twodash")+
  scale_x_continuous(breaks=c(1,2,3,4,5,6,7,8,9,10))

#cum lift + no model
ggplot(data=malm_d_avg_real, aes(x=p_rank, y=cumlift, group=1)) +
  geom_line(linetype = "dashed", color = "#69b3a2")+
  ggtitle('Cumulative lift chart')+
  geom_hline(yintercept=100, color = "black")+
  expand_limits(x = 1, y = 0)+
  scale_x_continuous(breaks=c(1,2,3,4,5,6,7,8,9,10))

#cum gains + no model
ggplot(data=malm_d_avg_real, aes(x=p_rank, y=cumgains, group=1)) +
  geom_line(linetype = "dashed", color = "#69b3a2")+
  ggtitle('Cumulative gains chart')+
  geom_abline(intercept = 0, slope = 0.1, color = "black")+
  expand_limits(x = 1, y = 0)+
  scale_x_continuous(breaks=c(1,2,3,4,5,6,7,8,9,10))
```

**Examining the effectiveness of the RFM analysis. Conduct the sequential RFM analysis** 

```{r}

#RFM
RFM <- select(malm, last,`total$`, purch, buyer)
RFM$buyer <- as.integer(as.logical(RFM$buyer)) # convert buyers into 1 true and 0 false

#monetary deciles
M <- RFM %>%
  mutate(
    m_rank = ntile(-`total$`, 10),     # Larger total value = more likely to buy
  )
M <- M %>% 
  group_by(m_rank) %>% 
  summarise(
    resprate = mean(buyer),
  )
ggplot(M, aes(x=m_rank, y=resprate)) + 
  geom_col(fill="#69b3a2", color="#e9ecef", alpha=0.9) + 
  ggtitle('Monetary Deciles')+
  scale_x_continuous(breaks=c(1,2,3,4,5,6,7,8,9,10))

#recency deciles
R <- RFM %>%
  mutate(
    r_rank = ntile(last, 10),     # Smaller recency value = more likely to buy
  )
R <- R %>% 
  group_by(r_rank) %>% 
  summarise(
    resprate = mean(buyer),
  )
ggplot(R, aes(x=r_rank, y=resprate)) + 
  geom_col(fill="#69b3a2", color="#e9ecef", alpha=0.9) + 
  ggtitle('Recency Deciles')+
  scale_x_continuous(breaks=c(1,2,3,4,5,6,7,8,9,10))
#frequency deciles
F <- RFM %>%
  mutate(
    f_rank = ntile(-purch, 10),     # Larger purch value = more likely to buy
  )
F <- F %>% 
  group_by(f_rank) %>% 
  summarise(
    resprate = mean(buyer),
  )
ggplot(F, aes(x=f_rank, y=resprate)) + 
  geom_col(fill="#69b3a2", color="#e9ecef", alpha=0.9) + 
  ggtitle('Frequency Deciles')+
  scale_x_continuous(breaks=c(1,2,3,4,5,6,7,8,9,10))
```

